{\begingroup
	\cleardoublepage
	\begin{acknowledgement}
		This thesis would not have been possible without the support of numerous individuals. First of all, I would like to thank Prof.\@~Ulrich Rückert for the unique opportunity to contribute to the Human Brain Project, and Thomas Schöpping and Dr.\@~Michael Thies for an introduction to the topic, guiding me along the right path and handing me valuable suggestions over the last year.
		
		I would also like to dedicate my thanks to the researchers at the Electronic Visions group at the Kirchhoff-Institute for Physics in Heidelberg, the Advanced Processor Technologies Research Group at the University of Manchester, and everyone else involved in the neuromorphic hardware and software platforms for answering my multitudinous questions and beavering away on the problems I encountered.

		On a technical side, none of this document would exist in this form without the work of thousands of volunteers who spent an astounding amount of time to contribute to the libre software projects used in writing and researching this thesis, including, but not limited to GCC, GNOME, the Linux kernel, Fedora, Octave, Matplotlib, Qt, \LaTeX, Gimp and Inkscape. Thank you for helping to create a free and better world.

		Furthermore, I wholeheartedly thank my parents for both the material and emotional support during my five-and-a-half-year exile in Bielefeld.% It is indeed a dull and rainy place (not as bad as England though).

		Last but not least, I express my sincere gratitude to all who read earlier revisions of this document and assisted me with their priceless feedback: Michael Thies, Thomas Schöpping, Benjamin Paaßen, Jan Philip Göpfert, Adriana Dreyer, Christoph Jenzen and Daniel Stöckel.
	\end{acknowledgement}
\endgroup}
